Hierarchy Builder: algebraic~hierarchies made easy in Coq with Elpi

### 2020-04-26 hb.tex
my 2 C

\Mixins{} are declared by the user as the fundamental building blocks of
a hierarchy. As the next proposition shows they shall not be distinguished
from regular \factories{}, since they are a special cases.

@@

\begin{command} to declare a new mixin:
\begin{coqcode}
#\hbmixin{}# Record M T of #$f_1\ \ldots\ f_n$# := { .. }.
\end{coqcode}
  Creates a module \coq{M} with a record which is exported as
  \coq{M}, and which depends on the \mixins{}
  \(\requires{}(\coq{M}) = \dep{}(\coq{M}) =
  \provides{}^\star\left(\enum{f_1}{f_n}\right)\), and registers this
  new record both as a \mixin{} and a \factory{}. It also exports an
  \phantterm{} \coq{M.Build.}
\end{command}

@@

  Definition #$b_{n+1}$# : #$f_{n+1}$# T .. := ...
  #\hbinstance{}# T #$b_{n+1}$#.
  ..
  Definition #$b_{n+k}$# : #$f_{n+k}$# T .. := ...
  #\hbinstance{}# T #$b_{n+k}$#.
#\hbend{}#.
\end{coqcode}
  Creates a module \coq{F} with a record which is exported as
  \coq{F}, and which depends on the mixins
  \(\requires{}(\coq{M}) =
  \provides{}^\star\left(\enum{f_1}{f_n}\right)\) and registers this
  new record as a factory.  It also exports an
  \phantterm{} \coq{F.Build.} We then use the \factoryinstances{}
  \(b_{n+1}, \ldots, b_{n+k}\) provided by the user and derive
  \mixinbuilders{} from them, so that
  \(\provides{}(\coq{F}) = \provides{}^\star\left(\enum{f_{n+1}}{f_{n+k}}\right).\)

  It is thus necessary that
  \(\requires{}^\star\left(\enum{f_{n+1}}{f_{n+k}}\right) \subseteq
  \provides{}^\star\left(\enum{f_1}{f_n}\right).\)
\end{commands}

Note that the \(b_i\) are not \mixinbuilders{} since their return types
are not necessarily \mixins{}, but could be \factories{}. However, since all
\factories{} provide \mixins{} through \mixinbuilders{}, we
obtain \mixinbuilders{} out of each \(b_i\) by function composition.

@@

\begin{definition}[\C{}, class]
A \class{} \coq{c} \(\in \C{}\) is distinguished a Coq record with one parameter
\coq{(T : Type)}. The type of each field is a \mixin{} in \M{} applied to
\coq{T} and, if needed, any number of other fields:
\begin{coqcode}
Record #$c$# (T : Type) := { #\(p_1 : m_1\ T\)#; .. ; #\(p_i : m_i\ T\ p_{\sigma(i,1)}\ \ldots p_{\sigma(i,q_i)}\)#; ..}.
\end{coqcode}
\end{definition}

>>>>

\Mixins{} are declared by the user as the fundamental building blocks of
a hierarchy. As the next proposition shows they shall not be regarded as
different from regular \factories{}, since they are a special cases.

@@

\begin{command} to declare a new mixin:
\begin{coqcode}
#\hbmixin{}# Record M T of #$f_1\ \ldots\ f_n$# := { .. }.
\end{coqcode}
  Declares the record \coq{M} inside an homonymous module \coq{M}.
  The record \coq{M.M} depends on the \mixins{}
  \(\requires{}(\coq{M}) = \dep{}(\coq{M}) =
  \provides{}^\star\left(\enum{f_1}{f_n}\right)\)
  and is registered both as a \mixin{} and a \factory{}. Finally It exports an
  \phantterm{} \coq{M.Build} and a notation
  \coq{M} standing for \coq{M.M}, so that the module name can be used to
  denote the homonymous record it contains.
\end{command}

@@

  Definition #$b_{n+1}$# : #$f_{n+1}$# T .. := ...
  #\hbinstance{}# T #$b_{n+1}$#.
  ..
  Definition #$b_{n+k}$# : #$f_{n+k}$# T .. := ...
  #\hbinstance{}# T #$b_{n+k}$#.
#\hbend{}#.
\end{coqcode}
  Declares the record \coq{F} inside an homonymous module \coq{F}.
  The record \coq{F.F} depends on the mixins
  \(\requires{}(\coq{M}) =
  \provides{}^\star\left(\enum{f_1}{f_n}\right)\) and is registered
  as a factory.  It exports an
  \phantterm{} \coq{F.Build.} and a notation
  \coq{F} standing for \coq{F.F}, so that the module name can be used to
  denote the homonymous record it contains.
  Finally it uses the \factoryinstances{}
  \(b_{n+1}, \ldots, b_{n+k}\) provided by the user in order to derive
  \mixinbuilders{}, so that
  \(\provides{}(\coq{F}) = \provides{}^\star\left(\enum{f_{n+1}}{f_{n+k}}\right).\)
  It is thus necessary that
  \(\requires{}^\star\left(\enum{f_{n+1}}{f_{n+k}}\right) \subseteq
  \provides{}^\star\left(\enum{f_1}{f_n}\right).\)
\end{commands}

Note that the \(b_i\) are not \mixinbuilders{} since their return types
are not necessarily \mixins{}, but could be \factories{}. However, since all
\factories{} provide \mixins{} through \mixinbuilders{}, we
can obtain \mixinbuilders{} out of each \(b_i\) by function composition.

@@

\begin{definition}[\C{}, class]
A \class{} \coq{c} \(\in \C{}\) is a distinguished Coq record with one parameter
\coq{(T : Type)}. The type of each field is a \mixin{} in \M{} applied to
\coq{T} and, if needed, any number of other fields:
\begin{coqcode}
Record #$c$# (T : Type) := { #\(p_1 : m_1\ T\)#; .. ; #\(p_i : m_i\ T\ p_{\sigma(i,1)}\ \ldots p_{\sigma(i,q_i)}\)#; ..}.
\end{coqcode}
\end{definition}
